Glue-on AdS holography for $T\bar T$-deformed CFTs

### 2023-06-21 alpha/GlueonAdS/glueon-beamer.tex
Update glueon slides

\documentclass[
	10pt
%	,draft
	,handout
	,noamsthm
]{beamer}

@@

\title{Glue-on AdS holography for $T\bar T$-deformed CFTs}
\subtitle{An extended AdS\,/\,\TTbar duality \textit{(beyond infinity)}}
\author{%
	Wen-Xin Lai 
	\texstringonly{\,}\textkai{赖文昕}%
	\texstringonly{
		\textit{\small in collaboration with} \\
			Luis Apolo,
			Peng-Xiang Hao \textkai{郝鹏翔}
			and Wei Song \textkai{宋伟}\\[2ex]
		\sidenote{\arxiv{2303.04836}}%
%		\ submitted to \textit{JHEP}
	}%
}
\institute{\small%
	Yau Mathematical Sciences Center,\\
	Tsinghua\texstringonly{
%		\\\large\ccbyncsajp
	}%
}
\date{}

@@

\begin{document}
{%%% TITLEPAGE
\setbeamercolor{title in head/foot}{
		use=palette quaternary
		,fg=palette quaternary.bg
	}
\logo{}
\begin{frame}
%	\titlegraphic{\vspace{2\baselineskip}}
	\titlepage
	\tikz [remember picture,overlay]
	\node at ([
		yshift=-4.8\baselineskip,
		xshift=.26\linewidth
	] current page.center) {
		\begin{minipage}{.5\textwidth}
		\centering
		\includegraphics[height=.33\textheight]{img/smbc-cantor-cropped.png}
		\\[.2ex]
		\footnotesize \url{https://smbc-comics.com/comic/cantor}
		\hspace{-1.5em}
		\end{minipage}
	};
\end{frame}
}%%% TITLEPAGE

@@

\section{\textbf{Warmup:} $\mrm{AdS}_{d+1}$/$\mrm{CFT}_{d}$ at $d = 2$}

@@

\begin{frame}{$\mrm{AdS}_3/\mrm{CFT}_2$ --- gravity trapped in a \textit{Poincar\'e cylinder}}{%
	\textcite{Aharony:1999ti}
}
\begin{columns}
\figAdsCft
\begin{column}{.75\textwidth}
	\begin{itemize}
	\item $\mrm{AdS}_3$ metric, $\rho \sim z^2 \sim r^{-2}$
	\begin{equation}
	\hspace{-2em}
		ds^2 = \ell^2 \frac{d\rho^2}{4 \rho^2} + \pqty{\frac{1}{\rho} g^{(0)}_{ij} + g^{(2)}_{ij} + \rho\thinmspace g^{(4)}_{ij} + \cdots}\, dx^i dx^j \label{fggauge}
	\end{equation}

@@

\section{\textbf{Review:} \TTbar deformation and cutoff AdS holography} \label{se:cutoffholography}

@@

\begin{frame}{\TTbar deformations --- ``solvable''}{%
	\citeTTbar
}
\begin{itemize}

@@

\begin{frame}{\TTbar deformations --- ``solvable''}{%
	\citeTTbar
}
\begin{itemize}

@@

\begin{frame}{Cutoff $\mrm{AdS}_3$ --- the dictionary}{%
	\textcite{McGough:2016lol} --
	\citetitle{McGough:2016lol}
}
\begin{columns}
\figCutoffAds
\hspace{-1em}
\begin{column}{.7\textwidth}
	\begin{itemize}
	\item Location (radius) of the cutoff surface:
	\begin{equation}
		\frac{1}{\zeta} \equiv \ell^{-2} g_{\varphi\varphi} \label{cutoff}
	\end{equation}
	gets mapped to the deformation parameter $\mu$:
	\begin{equation}
		\zeta_c = - \frac{c \mu}{3\ell^2}
		\label{dictionary}
	\end{equation}

@@

\begin{frame}{Cutoff $\mrm{AdS}_3$ --- gravity in a box}{%
	Title inspired by
	\textcite{Kraus:2021cwf} -- \citetitle{Kraus:2021cwf}%
}
\begin{columns}
\figCutoffAds
\hspace{-1em}
\begin{column}{.75\textwidth}
\vspace{-.5\baselineskip}
	\begin{itemize}
	\item \textbf{Significance:} gravity inside a finite box\\
		with hard Dirichlet walls

	\begin{itemize}
		\item Dual to some ``solvable'' non-conformal theory
		\item A step towards quantum gravity in reality!
	\end{itemize}

\pause
	\item \textbf{Caveat:} the duality only admits $\zeta_c > 0$ so $\mu < 0$
	\begin{equation}
		\zeta_c = - \frac{c \mu}{3\ell^2}
%		\label{dictionary}
	\end{equation}
	But \TTbar itself admits $\mu > 0$ with nice properties.\\
	\textit{What is the other side of the duality?}
	\end{itemize}
\end{column}
\end{columns}
\end{frame}

@@

\begin{frame}{Glue-on $\mrm{AdS}_3$ --- analytic continuation}{%
	\textcite{Apolo:2023vnm} -- \citetitle{Apolo:2023vnm}
}
	\begin{itemize}
	\item $\mrm{AdS}_3$ metric has only simple poles:
	\begin{equation}
		ds^2 = \ell^2 \bigg( \frac{d\rho^2}{4 \rho^2} + \frac{ \big( du + \rho \, \mathcal {\bar L}(v)\, dv \big) \big( dv + \rho \, \mathcal L(u)\, du \big) }{\rho} \bigg)\ %\label{banados}
	\end{equation}
	thus it admits well-defined analytic continuation from $\rho > 0$ to $\rho \in \mbb{R}$.
	\item Continuation of the dictionary:
	\begin{equation}
		\frac{1}{\zeta} \equiv \ell^{-2} g_{\varphi\varphi} \in \mbb{R},
	\ \quad
		\zeta_c = - \frac{c \mu}{3\ell^2} \in \mbb{R}
	\end{equation}
%	\vspace{-.5\baselineskip}

@@

\newcommand{\figGlueon}{
\begin{figure}[!h]
	\centering
	\includegraphics[width=.8\linewidth]{img/diagram.pdf}
	\vspace{-.3\baselineskip}
	\caption{Glue-on $\mrm{AdS}_3$}

	\vspace{-.3\baselineskip}
	\scriptsize\ Top-down view of the ``Poincare cylinder''
\end{figure}
}

\begin{frame}{Glue-on $\mrm{AdS}_3$ / \TTbar --- updating the dictionary}{%
	\textcite{Apolo:2023vnm} -- \citetitle{Apolo:2023vnm}
}
\begin{columns}
\begin{column}{.4\textwidth}
\figGlueon
\end{column}
\begin{column}{.6\textwidth}
\vspace{-\baselineskip}
\begin{equation}
	\frac{1}{\zeta} \equiv \ell^{-2} g_{\varphi\varphi} \in \mbb{R},
\ \quad
	\zeta_c = - \frac{c \mu}{3\ell^2} \in \mbb{R}
\end{equation}
\vspace{-.5\baselineskip}
\begin{itemize}
\item What have we done other than copy-pasting?
	\begin{itemize}
	\item Although the continuation is straightforward, physical quantities diverge as $\rho \to 0$

@@

\begin{frame}{Glue-on $\mrm{AdS}_3$ / \TTbar --- updating the dictionary}{%
	\textcite{Kraus:2018xrn} -- \citetitle{Kraus:2018xrn}
}
\begin{itemize}
\item Firstly, matching energy momentum \& the flow equations:
	\begin{equation}
		T_{ij}= \frac{\sigma_{\zeta_c}}{8\pi G} \Big( K_{ij} -  K \gamma_{ij} + \frac{1}{\ell |\zeta_c|}\gamma_{ij}\Big), \ \quad \sigma_{\zeta_c} = \frac{\abs{\zeta_c}}{\zeta_c} = \pm 1,
		\label{BYtensor}
	\end{equation}
	The field theory metric $\gamma_{ij} = \zeta_c h_{ij}$, where $h_{ij}$ is the induced metric.

@@

	\begin{itemize}
	\item $\gamma_{ij}$ is always positive-definite, while $h_{ij}$ becomes negative-definite for the glue-on region.\\
		This discrepancy is the origin of the absolute values.
	\end{itemize}

@@

\item \TTbar flow is recast geometrically as the $i,j$ components of the Einstein equations. \\
Note that the geometry can be decomposed such that:
	\begin{equation}
		ds^2 = \frac{1}{\zeta} \gamma_{ij}dx^i dx^j+n_\mu n_\nu dx^\mu dx^\nu
	\end{equation}
\end{itemize}
\end{frame}

@@

\pause
\item Treatment for the singularity at $\zeta\to 0^\pm$: introduce $\mathcal N_{\zeta={\pm\epsilon}}$ and glue them together\\
	exclude the $-\epsilon < \zeta < \epsilon$ region until finally sending $\epsilon \to 0$. 
\item We formally denote the boundary surface by the limit $\mathcal N_{0}=\mathcal N_{0^+}=\mathcal N_{0^-}$, though we need to keep track of the asymptotic cutoff $\epsilon$ in actual computations.
\end{itemize}
\end{frame}

\newcommand{\stateGlueon}{\ensuremath{
	\textrm{Cutoff / \textit{glue-on} $\mrm{AdS}_{d+1}$ Gravity}
	\ \equiv\ %
	\textrm{\TTbar deformed $\mrm{CFT}_{d}$ at $\mcal{N}_{\zeta_c}$}\qquad
}}

\begin{frame}{Glue-on $\mrm{AdS}_3$ / \TTbar --- the proposal}{%
	\textcite{Apolo:2023vnm} -- \citetitle{Apolo:2023vnm}
}
\centering
\vspace{-.5\baselineskip}
\begin{minipage}{.5\textwidth}
\centering
\includegraphics[width=.8\linewidth]{img/diagram.pdf}
\end{minipage}
\vspace{-.8\baselineskip}
\begin{equation}
\stateGlueon
\end{equation}
\end{frame}

@@

\begin{frame}{\TTbar quantities from bulk geometry}{%
	\textcite{Apolo:2023vnm} -- \citetitle{Apolo:2023vnm}
}
\begin{columns}
\begin{column}{.35\textwidth}
\figGlueon
\end{column}
\begin{column}{.65\textwidth}
\begin{itemize}
\item AdS/CFT: weak / strong duality
	\begin{itemize}
	\item \textit{quantum} quantities on the CFT side\\
		can be computed by (semi-)classical geometry
	\item inherited by the \TTbar deformation
	\end{itemize}

\pause
\item Demanding the extended geometry to be non-singular\\
	re-produces bounds on the \TTbar deformed theory, e.g.
	\begin{equation}
		\zeta_c \ge -1 \quad \Leftrightarrow \quad\mu\le \frac{ 3\ell^2 }{c} \label{reality}
	\end{equation}
	$\zeta_c = -1$ is a horizon of the glue-on geometry, \\
	where $\det g_{\mu\nu} = 0$.

@@

\begin{frame}{\TTbar spectrum from bulk covariant charges}{%
	\textcite{Kraus:2021cwf,Apolo:2023vnm}
}
\begin{columns}
\begin{column}{.35\textwidth}
\figGlueon
\end{column}
\begin{column}{.65\textwidth}
\vspace{-.3\baselineskip}
\begin{itemize}
\item Spectrum: conserved charges of geometries,\\
	e.g.~black holes \textit{BTZ} \cite{Banados:1992wn}
	\begin{itemize}
	\item computed with the covariant formalism\\
		\textcite{Iyer:1994ys, Barnich:2001jy}
	\item manifest covariance (diff-invariance) depends on reparametrization redundancies (gauge),\\
		so delay gauge fixing until the very end
	\end{itemize}

@@

\item This reproduces $E(\mu)$, $J(\mu)$ with $\mu \in\mbb{R}$ in \eqref{ttbarspectrum}.

@@

\begin{frame}{\TTbar partition functions from bulk on-shell actions}{%
	\textcite{Kraus:2021cwf,Apolo:2023vnm}
}
\begin{columns}
\begin{column}{.35\textwidth}
\figGlueon
\end{column}
\begin{column}{.65\textwidth}
\vspace{-.3\baselineskip}
\begin{itemize}
\item Bulk: weakly coupled gravity, the partition function \eqref{eq:partition} is approximated by the on-shell gravity action of the dominant saddle (method of steepest descent):
	\begin{equation}
		Z_{T\bar T} (\mu) = \mathcal Z (\zeta_c) \approx  e^{-I(\zeta_c)}\label{partition2}
	\end{equation}

@@

@@

\begin{frame}{Glue-on $\mrm{AdS}_3$ / \TTbar deformed $\mrm{CFT}_2$}{%
	\textcite{Apolo:2023vnm} -- \citetitle{Apolo:2023vnm}
}
\vspace{-1.2\baselineskip}
\begin{equation}
\stateGlueon
\end{equation}
\vspace{-.3\baselineskip}
\begin{columns}
\begin{column}{.3\textwidth}
\centering
\includegraphics[width=\linewidth]{img/diagram.pdf}
\end{column}
\hspace{-1em}
\begin{column}{.7\textwidth}
\begin{itemize}
\item Holographic proposal for $T\bar T$-deformed CFTs with $\mu \in \mbb{R}$

@@

>>>>

\documentclass[
	10pt
%	,draft
%	,handout
	,noamsthm
]{beamer}

@@

\title{Glue-on AdS holography for $T\bar T$-deformed CFTs}
\subtitle{An extended AdS\,/\,\TTbar duality \textit{(beyond infinity)}}
\author{%
	Wen-Xin Lai 
	\texstringonly{\,}\textkai{赖文昕}%
	\texstringonly{
		\textit{\small in collaboration with} \\
			Luis Apolo,
			Peng-Xiang Hao \textkai{郝鹏翔}
			and Wei Song \textkai{宋伟}\\[2ex]
		\sidenote{\arxiv{2303.04836}}%
%		\ submitted to \textit{JHEP}
	}%
}
\institute{\small%
	Yau Mathematical Sciences Center,\\
	Tsinghua\texstringonly{
%		\\\large\ccbyncsajp
	}\\[2ex]
	\today\ @ BIMSA
}
\date{}

@@

\begin{document}
{%%% TITLEPAGE
\setbeamercolor{title in head/foot}{
		use=palette quaternary
		,fg=palette quaternary.bg
	}
\logo{}
\begin{frame}
%	\titlegraphic{\vspace{2\baselineskip}}
	\titlepage
	\tikz [remember picture,overlay]
	\node at ([
		yshift=-4.6\baselineskip,
		xshift=.25\linewidth
	] current page.center) {
		\begin{minipage}{.5\textwidth}
		\centering
		\includegraphics[height=.33\textheight]{img/smbc-cantor-cropped.png}
		\\[.2ex]
		\footnotesize \url{https://smbc-comics.com/comic/cantor}
		\hspace{-1.5em}
		\end{minipage}
	};
\end{frame}
}%%% TITLEPAGE

@@

\section{\textbf{Warmup:} Pure Einstein gravity in $\mrm{AdS}_3$ / $\mrm{CFT}_2$}

@@

\begin{frame}{$\mrm{AdS}_3/\mrm{CFT}_2$ --- gravity trapped in a cylinder}{%
	\textcite{Aharony:1999ti}
}
\begin{columns}
\figAdsCft
\begin{column}{.75\textwidth}
	\begin{itemize}
	\item $\mrm{AdS}_3$ metric, $\rho \sim z^2 \sim r^{-2}$
	\begin{equation}
	\hspace{-2em}
		ds^2 = \ell^2 \frac{d\rho^2}{4 \rho^2} + \pqty{\frac{1}{\rho} g^{(0)}_{ij} + g^{(2)}_{ij} + \rho\thinmspace g^{(4)}_{ij} + \cdots}\, dx^i dx^j \label{fggauge}
	\end{equation}

@@

\section{\textbf{Review:} \TTbar deformation and cutoff AdS} \label{se:cutoffholography}

@@

\begin{frame}{\TTbar deformations --- definition}{%
	\citeTTbar
}
\begin{itemize}

@@

\begin{frame}{``Solvable'' deformations on the field theory side}{%
	\citeTTbar
}
\begin{itemize}

@@

\begin{frame}{Cutoff $\mrm{AdS}_3$ --- the dictionary}{%
	\textcite{McGough:2016lol} --
	\citetitle{McGough:2016lol}
}
\begin{columns}
\figCutoffAds
\hspace{-1em}
\begin{column}{.7\textwidth}
	\begin{itemize}
	\item Location (radius) of the cutoff surface:
	\begin{equation}
		\frac{1}{\zeta} \equiv \ell^{-2} g_{\varphi\varphi} \label{cutoff}
	\end{equation}
	gets mapped to the deformation parameter $\mu$:
	\begin{equation}
		\zeta_c = - \frac{c \mu}{3\ell^2}
		\label{dictionary}
	\end{equation}
	\textit{Roughly $\mu \sim 1/r^2$}

@@

\begin{frame}{Cutoff $\mrm{AdS}_3$ --- gravity in a box}{%
	Title inspired by
	\textcite{Kraus:2021cwf} -- \citetitle{Kraus:2021cwf}%
}
\begin{columns}
\figCutoffAds
\hspace{-1.2em}
\begin{column}{.75\textwidth}
\vspace{-.5\baselineskip}
	\begin{itemize}
	\item \textbf{Significance:} gravity inside a finite box\\
		with hard Dirichlet walls

	\begin{itemize}
		\item Dual to some ``solvable'' no-longer-conformal ft
		\item A step towards quantum gravity in reality!
	\end{itemize}

\pause
	\item \textbf{Caveat:} the duality only admits $\zeta_c > 0$ so $\mu < 0$
	\begin{equation}
		\zeta_c = - \frac{c \mu}{3\ell^2}
%		\label{dictionary}
	\end{equation}
	But \TTbar itself admits $\mu > 0$ with nice properties.\\
	\textit{What is the other side of the duality?}
		\begin{itemize}
		\item For comparison, the proposal of \textcite{Guica:2019nzm} admits both signs of $\mu$
		\end{itemize}
	\end{itemize}
\end{column}
\end{columns}
\end{frame}

@@

\begin{frame}{Glue-on $\mrm{AdS}_3$ --- analytic continuation}{%
	\textcite{Apolo:2023vnm} -- \citetitle{Apolo:2023vnm}
}
	\begin{itemize}
	\item $\mrm{AdS}_3$ metric has only simple poles with the $\rho$ coordinate:
	\begin{equation}
		ds^2 = \ell^2 \bigg( \frac{d\rho^2}{4 \rho^2} + \frac{ \big( du + \rho \, \mathcal {\bar L}(v)\, dv \big) \big( dv + \rho \, \mathcal L(u)\, du \big) }{\rho} \bigg)\ %\label{banados}
	\end{equation}
	thus it admits well-defined analytic continuation \\
	from $\rho > 0$ to $\rho \in \mbb{R}$.
	\item Continuation of the dictionary:
	\begin{equation}
		\frac{1}{\zeta} \equiv \ell^{-2} g_{\varphi\varphi} \in \mbb{R},
	\ \quad
		\zeta_c = - \frac{c \mu}{3\ell^2} \in \mbb{R}
	\end{equation}
%	\vspace{-.5\baselineskip}

@@

\newcommand{\figGlueon}[1][.37]{
\begin{column}{#1\textwidth}
\begin{figure}[!h]
	\centering
	\includegraphics[width=\linewidth]{img/diagram.pdf}
	\vspace{-1.5\baselineskip}
	\caption{Glue-on $\mrm{AdS}_3$}

	\vspace{-.3\baselineskip}
	\scriptsize\ Top-down view of the Poincar\'e disk
\end{figure}
\end{column}
}

\begin{frame}{Glue-on $\mrm{AdS}_3$ / \TTbar --- updating the dictionary}{%
	\textcite{Apolo:2023vnm} -- \citetitle{Apolo:2023vnm}
}
\begin{columns}
\figGlueon
\begin{column}{.64\textwidth}
\vspace{-\baselineskip}
\begin{equation}
	\frac{1}{\zeta} \equiv \ell^{-2} g_{\varphi\varphi} \in \mbb{R},
\ \quad
	\zeta_c = - \frac{c \mu}{3\ell^2} \in \mbb{R}
\end{equation}
\vspace{-.5\baselineskip}
\begin{itemize}
\item What have we done other than copy-pasting?
	\begin{itemize}
	\item Although the continuation is straightforward, physical quantities diverge as $\rho \to 0$

@@

\begin{frame}{Glue-on $\mrm{AdS}_3$ / \TTbar --- updating the dictionary}{%
	\textcite{Kraus:2018xrn} -- \citetitle{Kraus:2018xrn}
}
\begin{itemize}
\item Firstly, matching energy momentum (Brown-York) \& the flow equations:
	\begin{equation}
		T_{ij}= \frac{\sigma_{\zeta_c}}{8\pi G} \Big( K_{ij} -  K \gamma_{ij} + \frac{1}{\ell |\zeta_c|}\gamma_{ij}\Big), \ \quad \sigma_{\zeta_c} = \frac{\abs{\zeta_c}}{\zeta_c} = \pm 1,
		\label{BYtensor}
	\end{equation}
	The field theory metric $\gamma_{ij} = \zeta_c h_{ij}$, where $h_{ij}$ is the induced metric.

@@

	\begin{itemize}
	\item $\gamma_{ij}$ is always positive-definite, while $h_{ij}$ becomes negative-definite for the glue-on region.
	This discrepancy is the origin of the $\abs{\zeta_c}$.
	\end{itemize}

@@

\item \TTbar flow is recast geometrically as the $i,j$ components of the Einstein equations. 
Note that the geometry can be decomposed such that:
	\begin{equation}
		ds^2 = \frac{1}{\zeta} \gamma_{ij}dx^i dx^j+n_\mu n_\nu dx^\mu dx^\nu
	\end{equation}
\end{itemize}
\end{frame}

@@

\pause
\item Treatment for the singularity at $\zeta\to 0^\pm$: introduce $\mathcal N_{\zeta={\pm\epsilon}}$ and glue them together (``glue-on'');
	exclude the $-\epsilon < \zeta < \epsilon$ region until finally sending $\epsilon \to 0$. 
\item We formally denote the boundary surface by the limit $\mathcal N_{0}=\mathcal N_{0^+}=\mathcal N_{0^-}$, though we need to keep track of the asymptotic cutoff $\epsilon$ in actual computations.
\end{itemize}
\end{frame}

\newcommand{\stateGlueon}{\ensuremath{
	\textrm{Cutoff / \textit{glue-on} $\mrm{AdS}_{d+1}$ Gravity}
	\ \equiv\ %
	\textrm{\TTbar deformed $\mrm{CFT}_{d}$ at $\mcal{N}_{\zeta_c}$}\ \,
}}

\begin{frame}{Glue-on $\mrm{AdS}_3$ / \TTbar --- the proposal}{%
	\textcite{Apolo:2023vnm} -- \citetitle{Apolo:2023vnm}
}
\centering
\vspace{-.5\baselineskip}
\begin{minipage}{.5\textwidth}
\centering
\includegraphics[width=.9\linewidth]{img/diagram.pdf}
\end{minipage}
\vspace{-.1\baselineskip}
\begin{equation*}
\stateGlueon
\end{equation*}
\end{frame}

@@

\begin{frame}{\TTbar quantities from bulk geometry}{%
	\textcite{Apolo:2023vnm} -- \citetitle{Apolo:2023vnm}
}
\begin{columns}
\figGlueon[.3]
\begin{column}{.75\textwidth}
\begin{itemize}
\item AdS/CFT: weak / strong duality
	\begin{itemize}
	\item \textit{quantum} quantities on the CFT side\\
		can be computed by (semi-)classical geometry
	\item inherited by the \TTbar deformation
	\end{itemize}

\pause
\item Demanding the extended geometry to be non-singular
	re-produces bounds on the \TTbar deformed theory, e.g.
	\begin{equation}
		\zeta_c \ge -1 \quad \Leftrightarrow \quad\mu\le \frac{ 3\ell^2 }{c} \label{reality}
	\end{equation}
	$\zeta_c = -1$ is a horizon of the glue-on geometry, \\
	where $\det g_{\mu\nu} = 0$.

@@

\begin{frame}{\TTbar spectrum from bulk covariant charges}{%
	\textcite{Kraus:2021cwf,Apolo:2023vnm}
}
\begin{columns}
\figGlueon[.3]
\begin{column}{.75\textwidth}
\vspace{-.3\baselineskip}
\begin{itemize}
\item Spectrum: conserved charges of geometries,\\
	e.g.~black holes \textit{BTZ} \cite{Banados:1992wn}
	\begin{itemize}
	\item computed with the covariant formalism\\
		\textcite{Iyer:1994ys, Barnich:2001jy}
	\item manifest covariance (diff-invariance) depends on reparametrization redundancies (gauge),\\
		so delay gauge fixing until the very end
	\end{itemize}

@@

\item This reproduces $E(\mu)$, $J(\mu)$ with $\mu \in\mbb{R}$ in \eqref{ttbarspectrum}

@@

\begin{frame}{\TTbar spectrum from bulk covariant charges}{%
	\textcite{Kraus:2021cwf,Apolo:2023vnm}
}
\begin{columns}
%\figGlueon[.3]
\begin{column}{\textwidth}
\vspace{-.3\baselineskip}
\begin{itemize}
\item Covariant charges of geometries
	\begin{align*}
		\delta E(\mu) &\equiv  \ell^{-1} \delta \mathcal Q_{\pd_{t'}}, \\ \delta J(\mu) &\equiv  \delta \mathcal Q_{\pd_{\varphi'}},
	\end{align*}

\item One needs to choose the appropriate boundary coordinates $(t',\varphi')$, such that
	\begin{equation}
		ds^2_c = \ell^2 ( -dt'^2 + d\varphi'^2) , \qquad \varphi' \sim \varphi' + 2\pi.\label{cutoffmetric}
	\end{equation}

\item This is realized by the \textit{state-dependent} map of coordinates
	\begin{align*}
		dt' &= \sqrt{ \big(1 - \zeta_c (T_u+T_v)^2 \big) \big(1 - \zeta_c (T_u-T_v)^2 \big) } \, dt,  \\
		 d\varphi' &= d\varphi + \zeta_c \big( T_u^2 - T_v^2 \big)\,dt.
	\end{align*}
\end{itemize}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{\TTbar thermodynamics from bulk geometry}{%
	\textcite{Giveon:2017nie,Apolo:2019zai}
}
\begin{columns}
%\figGlueon[.3]
\begin{column}{\textwidth}
\vspace{-.3\baselineskip}
\begin{itemize}

\item \textit{State-dependent} map of coordinates
	\begin{align*}
		dt' &= \sqrt{ \big(1 - \zeta_c (T_u+T_v)^2 \big) \big(1 - \zeta_c (T_u-T_v)^2 \big) } \, dt, \\
		 d\varphi' &= d\varphi + \zeta_c \big( T_u^2 - T_v^2 \big)\,dt.
	\end{align*}

\pause
\item Many consequences: correct charges, signal propagation speed $v'_{\pm} \equiv \pm \frac{d\varphi'}{dt'}$, and thermodynamics when Wick rotated:
	\begin{equation}
		T_L(\mu)\,T_R(\mu) \le - \frac{1}{4\pi^2 \ell^2 \zeta_c} = \frac{3}{4\pi^2c\mu} = T_H(\mu)^2,
	\quad \mu > 0
	\end{equation}
	$T_{L,R}$ are temperatures associated with $u',v' = \varphi' \pm t'$.

\item $T_H$ is the \textit{Hagedorn temperature}: exceeding $T_H$ corresponds to a complex entropy
\end{itemize}
\end{column}
\end{columns}
\end{frame}

\section{\textbf{Check:} \TTbar partition functions from glue-on AdS} \label{se:partitionfunction}

\begin{frame}{\TTbar partition functions from bulk on-shell actions}{%
	\textcite{Caputa:2020lpa}
}
\begin{columns}
\figGlueon
\hspace{-1.5em}
\begin{column}{.7\textwidth}
\vspace{-.3\baselineskip}
\begin{itemize}
\item Bulk: weakly coupled gravity, the partition function is approximated by the on-shell gravity action of the dominant saddle:
	\begin{equation}
		Z_{T\bar T} (\mu) = \mathcal Z (\zeta_c) \approx  e^{-I(\zeta_c)}\label{partition2}
	\end{equation}

@@

\begin{frame}{Comparison with field theory analysis}{%
	\textcite{Datta:2018thy,Apolo:2023aho}
}
\begin{columns}
%\figGlueon
%\hspace{-1.5em}
\begin{column}{\textwidth}
\vspace{-.3\baselineskip}
\begin{itemize}
\item \textbf{Torus:} Large $c$, \textit{modular invariance} \& sparseness of the ``light'' spectrum \cite{Datta:2018thy,Apolo:2023aho},
	c.f. \textcite{Hartman:2014oaa}
	\begin{equation*}
	\hspace{-1.5em}
		\log   Z_{T\bar T}(\mu)  \approx \left\{ \begin{aligned}
		& {-\frac{1}{2}}\,(\beta_L + \beta_R)\, RE_{\text{vac}}(\mu),  &\beta_L \beta_R > 1, \\
		& {-2 \pi^2 \bigg(\frac{1}{\beta_L} + \frac{1}{\beta_R}\bigg)  RE_{\text{vac}}\bigg(\frac{4\pi^2}{\beta_L \beta_R} \mu \bigg)}, \ \, &\beta_L\beta_R < 1, %\\[2ex]
		 \end{aligned} \right. %\label{ZTTbar}
	\end{equation*}

\item \textbf{Sphere:} maximally symmetric, \textcite{Donnelly:2018bef}
	\begin{equation}
%		\label{sphere:stresstensor}
		\langle T_{ij}\rangle =-\frac{1}{4\pi\mu} \bigg(1-\sqrt{1-\frac{c\mu}{3L^2}}\, \bigg)  \gamma_{ij}.
	\end{equation}
\end{itemize}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{Sphere partition functions}{%
	\textcite{Donnelly:2018bef,Li:2020zjb}
}
\begin{columns}
%\figGlueon
%\hspace{-1.5em}
\begin{column}{\textwidth}
\vspace{-.3\baselineskip}
\begin{itemize}

\item Given the explicit stress tensor $\langle T_{ij}\rangle \propto \gamma_{ij}$, the flow equations:
	\begin{align*}
\partial_\mu \log Z_{\TTbar}(\mu)&=8\pi \int d^2x\sqrt{\gamma}\, \langle T\bar{T} \rangle \\
-L\,\partial_L \log Z_{\TTbar}(\mu)&=\int d^2x\sqrt{\gamma}\,\langle T^i_i \rangle
	\end{align*}

	\item ... admit the general solution with a $\mu$-\textit{independent} integration $a$:
	\begin{equation*}
		\log Z_{\TTbar}(\mu, a) = \frac{c}{3} \log \bigg[\frac{L}{a}   \bigg(1+\sqrt{1-\frac{c \mu }{3  L^2}}\, \bigg) \bigg] - \frac{L^2}{\mu}  \sqrt{1-\frac{c \mu }{3 L^2}} + \frac{L^2}{\mu} %\label{Zsol}
	\end{equation*}

	\item \textcite{Donnelly:2018bef} is recovered with $a = \sqrt{c|\mu|/3}$ but one can also take $a = \epsilon$, thus decoupling the energy (RG) scale $\epsilon$ with the deformation $\mu$. 
\end{itemize}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{Gravitational actions: with counterterms}{%
	\textcite{Donnelly:2018bef,Li:2020zjb}
}
\begin{columns}
%\figGlueon
%\hspace{-1.5em}
\begin{column}{1.1\textwidth}
\vspace{-.3\baselineskip}
\begin{itemize}

\item Gravitational actions, with counterterms:
	\begin{align*}
		I_{\mcal{M}}(\zeta_1, \zeta_2) & \coloneqq - \frac{1}{16\pi G} \int_{\zeta_1}^{\zeta_2} d\zeta \int d^2 x \sqrt{g} \, (R + 2 \ell^{-2}),  %\label{bulkaction} \\
	\\
		I_{\mathcal N_{\zeta}} & \coloneqq  - \frac{1}{8\pi G} \int_{\mathcal N_{\zeta}} d^2x \sqrt{h}\,h^{ij} K_{ij} 
	\\& \qquad + \frac{\sigma_\zeta}{8\pi G} \int_{\mathcal N_{\zeta}} d^2x \sqrt{h} \, \bigg(\ell^{-1} - \frac{\ell  \mathcal{R}[h]}{4} \log |\zeta| \bigg), %\label{boundaryaction}
	\end{align*}
	... consistent with the Brown-York stress tensor $T_{ij}$, \\
	fully renormalized at $\zeta \to 0^\pm$.

	\item The lack of diff invariance of the $\log |\zeta|$ counterterm is known to be a reflection of the Weyl anomaly: \textsl{\citeauthor{Henningson:1998gx,deHaro:2000vlm,Papadimitriou:2010as}}.

\begin{frame}{Gravitational actions: with counterterms}{%
	\textcite{Caputa:2020lpa,Li:2020zjb}
}
\begin{columns}
%\figGlueon
%\hspace{-1.5em}
\begin{column}{1.1\textwidth}
\vspace{-.3\baselineskip}
\begin{itemize}

\item The $\log$ counterterm makes a crucial contribution to the on-shell action:
	\begin{align*}
		I_{\mcal{M}}(\zeta_1, \zeta_2) & \coloneqq - \frac{1}{16\pi G} \int_{\zeta_1}^{\zeta_2} d\zeta \int d^2 x \sqrt{g} \, (R + 2 \ell^{-2}),  %\label{bulkaction} \\
	\\
		I_{\mathcal N_{\zeta}} & \coloneqq  - \frac{1}{8\pi G} \int_{\mathcal N_{\zeta}} d^2x \sqrt{h}\,h^{ij} K_{ij} 
	\\& \qquad + \frac{\sigma_\zeta}{8\pi G} \int_{\mathcal N_{\zeta}} d^2x \sqrt{h} \, \bigg(\ell^{-1} - \frac{\ell  \mathcal{R}[h]}{4} \log |\zeta| \bigg), %\label{boundaryaction}
	\end{align*}
	It guarantees that the partition function is compatible with the $T\bar T$ differential equation --- not the case for \textsl{\citeauthor{Donnelly:2018bef}}.

	\item In general the space of deformed theories are parametrized by $(\mu,a)$, where $a$ is the length scale (inverse energy scale). In many contexts they are naturally related: $a = \sqrt{c|\mu|/3}$, but $a$ can be tuned by RG.

@@

\begin{frame}{Glue-on $\mrm{AdS}_3$ / \TTbar deformed $\mrm{CFT}_2$}{%
	\textcite{Apolo:2023vnm} -- \citetitle{Apolo:2023vnm}
}
\vspace{-2\baselineskip}
\begin{equation*}
\stateGlueon
\end{equation*}
\vspace{-1.3\baselineskip}
\begin{columns}
\begin{column}{.37\textwidth}
\begin{figure}[!h]
	\centering
	\includegraphics[width=\linewidth]{img/RT-AdS.pdf}
	\vspace{-1.5\baselineskip}
	\caption{RT for Glue-on $\mrm{AdS}_3$}

	\vspace{-.3\baselineskip}
	\scriptsize\, Top-down view of the Poincar\'e disk
\end{figure}
\end{column}
\hspace{-1em}
\begin{column}{.71\textwidth}
\begin{itemize}
\item Holographic proposal for $T\bar T$-deformed CFTs\\
with $\mu \in \mbb{R}$.

@@

\begin{frame}{Glue-on $\mrm{AdS}_3$ / \TTbar deformed $\mrm{CFT}_2$}{%
	\textcite{Apolo:2023vnm} -- \citetitle{Apolo:2023vnm}
}
\vspace{-2\baselineskip}
\begin{equation*}
\stateGlueon
\end{equation*}
\vspace{-1.3\baselineskip}
\begin{columns}
\begin{column}{.37\textwidth}
\begin{figure}[!h]
	\centering
	\includegraphics[width=\linewidth]{img/RT-AdS.pdf}
	\vspace{-1.5\baselineskip}
	\caption{RT for Glue-on $\mrm{AdS}_3$}

	\vspace{-.3\baselineskip}
	\scriptsize\, Top-down view of the Poincar\'e disk
\end{figure}
\end{column}
\hspace{-1em}
\begin{column}{.71\textwidth}
\begin{itemize}
